---
title: "Pool Together Analysis"
author: "Omni Analytics Group"
date: "8/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)

library(tidyverse)
library(knitr)
```

## Introduction

PoolTogether is a fun, open sourced and decentralized protocol for “no-loss” lotteries where users transfer their crypto into a shared pool that acts as a “prize-linked savings account”. These deposits double as lottery tickets that give the depositor a chance to win proportional to the amount of the pool they claim. Just like traditional savings accounts, all money in the pool earns a yield and each user can withdraw at their discretion.

At the time of this writing, \$3,801,856 worth of prizes have been awarded. Also, based on current deposits, the platform has been generating \$93,558 in interest per week for its users. Interested in earning? You can participate in PoolTogether on the Ethereum, xDai, Matic, and Binance Smart Chain networks.

True to the spirit of decentralization, the PoolTogether protocol is managed by holders of the $POOL governance token who draft and vote on proposals that control the prize creation tools and prize pools. Proposals are ratified through a voting process and are then implemented by the community.

## Data

To begin, we need to scrape the data. The code is provided below, if desired. The procedure for data scraping is:

1. Retrieve the network/chain pool data from the pool together API json files.
2. Parse the JSON / nested list structure into an R dataframe.
3. Retrieve the rewards / prize history by making a GraphQL call to TheGraph's API.
4. Retrieve information on the holders by making a GraphQL call to TheGraph's API.
5. Save the three objects as R Data Files.

<details><summary>View Code</summary>

```{r code, results='hide', eval=FALSE}
## Load Required libraries
library(httr)
library(jsonlite)
library(lubridate)
library(ghql)

## Fetch data
chain1_data <- GET("https://pooltogether-api.com/pools/1.json")
chain137_data <- GET("https://pooltogether-api.com/pools/137.json")
parsed_pools <- c(content(chain1_data,"parsed"),content(chain137_data,"parsed"))

lapply(parsed_pools,function(x) names(x$tokens))

## Parse Pool
null_to_na <- function(x) ifelse(is.null(x),NA,x)
parse_pool <- function(pool)
{
	data.frame(	
				Name = pool$name,
				Symbol = pool$symbol,
				Address = pool$contract$prizePool$address,
				Subgraph_Version = pool$contract$subgraphVersion,
				Num_Winners = pool$config$numberOfWinners,
				Prize_Freq_Secs = pool$config$prizePeriodSeconds,
				Pool_Type = pool$prizePool$type,

				Base_Tok_Address = pool$tokens$underlyingToken$address,
				Base_Tok_Decimals = pool$tokens$underlyingToken$decimals,
				Base_Tok_Name = pool$tokens$underlyingToken$name,
				Base_Tok_Symbol = pool$tokens$underlyingToken$symbol,
				Base_Tok_Val_USD = pool$tokens$underlyingToken$usd,
				
				Ticket_Address = pool$tokens$ticket$address,
				Ticket_Decimals = pool$tokens$ticket$decimals,
				Ticket_Name = pool$tokens$ticket$name,
				Ticket_Symbol = pool$tokens$ticket$symbol,
				Num_Ticket_Holders = pool$tokens$ticket$numberOfHolders,
				Ticket_Val_Base_Tok = pool$tokens$ticket$totalSupply,
				Ticket_Val_USD = pool$tokens$ticket$totalValueUsd,

				Sponsorship_Address = pool$tokens$sponsorship$address,
				Sponsorship_Decimals = pool$tokens$sponsorship$decimals,
				Sponsorship_Name = pool$tokens$sponsorship$name,
				Sponsorship_Symbol = pool$tokens$sponsorship$symbol,
				Num_Sponsorship_Holders = pool$tokens$sponsorship$numberOfHolders,
				Sponsorship_Val_Base_Tok = pool$tokens$sponsorship$totalSupply,
				Sponsorship_Val_USD = pool$tokens$sponsorship$totalValueUsd,

				Reserve_Val_Base_Tok = pool$reserve$amount,
				Reserve_Val_USD = pool$reserve$totalValueUsd,
				Reserve_Rate = pool$reserve$rate,

				Daily_Reward = null_to_na(pool$tokenListener$dripRatePerDay),
				Daily_Reward_Token_Name = null_to_na(pool$tokens$tokenFaucetDripToken$name),
				Daily_Reward_Token_Decimals = null_to_na(pool$tokens$tokenFaucetDripToken$decimals),
				Daily_Reward_Token_Symbol = null_to_na(pool$tokens$tokenFaucetDripToken$symbol),
				Daily_Reward_Token_Val_USD = null_to_na(pool$tokens$tokenFaucetDripToken$usd),
				Effective_APR = null_to_na(pool$tokenListener$apr),
 
				Prize_Current_Index = pool$prize$currentPrizeId,
				Prize_Current_State = pool$prize$currentState,
				Prize_Remaining_Secs = strtoi(pool$prize$prizePeriodRemainingSeconds$hex),
				Prize_Period_Started_At = as_datetime(strtoi(pool$prize$prizePeriodStartedAt$hex)),
 

				Prize_Total_USD = pool$prize$totalValueUsd,
				Prize_Internal = pool$prize$totalInternalAwardsUsd,
				Prize_External = pool$prize$totalExternalAwardsValueUsd,
				Prize_External_Yield = null_to_na(pool$prize$yield$comp$totalValueUsd),
				Prize_External_ERC20 = pool$prize$erc20Awards$totalValueUsd

			)
}
pools_data <- do.call(rbind,lapply(parsed_pools,parse_pool))
saveRDS(pools_data,"data/pools_data.RDS")


####################################################
## Collect Past Prize History
####################################################
get_rewards <- function(poolAddress,version)
{
	grph_url <- paste0("https://api.thegraph.com/subgraphs/name/pooltogether/pooltogether-v",gsub("\\.","_",version))
	con_r <- GraphqlClient$new(grph_url)
	qry_r <- Query$new()
	qry_r$query('prize_data',
		'query prize_data($poolAddress: String!)
		{
			prizePool(id:$poolAddress)
			{
				prizes(first:1000,orderBy: awardedTimestamp, orderDirection: desc)
				{
					id
					awardedTimestamp
					awardedBlock
					totalTicketSupply
					prizePeriodStartedTimestamp
					lockBlock
					awardedControlledTokens
					{
						id
						amount
						token
						winner
						prize
						{
							id
							prizePool
							{
								underlyingCollateralDecimals
							}
						}
					}
					awardedExternalErc20Tokens
					{
						id
						winner
						address
						balanceAwarded
						name
						symbol
						decimals
					}
					awardedExternalErc721Nfts
					{
						id
						address
						tokenIds
						winner
					}
				}
			}
		}')
	rewards <- fromJSON(con_r$exec(qry_r$queries$prize_data, list(poolAddress = poolAddress)))$data$prizePool$prizes
	rewards_df <- data.frame(
								Awarded_Block = rewards$awardedBlock,
								Awarded_Time = as_datetime(as.numeric(rewards$awardedTimestamp)),
								Prize_Period_Started_At = as_datetime(as.numeric(rewards$prizePeriodStartedTimestamp)),
								Ticket_Supply = rewards$totalTicketSupply,
								Prize_External_ERC20 = do.call(c,lapply(rewards$awardedExternalErc20Tokens,function(x) paste0("[",x$address,";",as.numeric(x$balanceAwarded)/(10^as.numeric(x$decimals)),";",x$name,";",x$symbol,"]",collapse=""))),
								Prize_Internal = do.call(c,lapply(rewards$awardedControlledTokens,function(x) sum(as.numeric(x$amount)/(10^as.numeric(x$prize$prizePool$underlyingCollateralDecimals)))))
							)
	return(rewards_df)

}
rewards_data <- list()
for(idx in 1:nrow(pools_data))
{
	rewards_data[[idx]] <- get_rewards(pools_data$Address[idx],pools_data$Subgraph_Version[idx])
	message(idx)
}
saveRDS(rewards_data,"data/rewards_data.RDS")
####################################################
####################################################

####################################################
## Collect Holder History
####################################################
get_holders <- function(ticketAddress,version)
{
	grph_url <- paste0("https://api.thegraph.com/subgraphs/name/pooltogether/pooltogether-v",gsub("\\.","_",version))
	con_h <- GraphqlClient$new(grph_url)
	qry_h <- Query$new()
	qry_h$query('holder_data',
	'query holder_data($ticketAddress: String!,$first: Int!, $skip: Int!)
	{
		controlledTokenBalances(first:$first,skip:$skip,orderBy: balance, orderDirection: desc,where:{controlledToken:$ticketAddress})
		{
			balance
			account
			{
				id
			}
		}
	}')
	first = 1000
	skip = 0
	holders = data.frame()
	while(TRUE)
	{
		holders_t <- fromJSON(con_h$exec(qry_h$queries$holder_data, list(ticketAddress = ticketAddress,first=first,skip=skip)))$data$controlledTokenBalances
		holders_t <- data.frame(id = holders_t$account$id,balance = holders_t$balance)
		if(nrow(holders_t) == 0) break()
		rownames(holders_t) <- as.character(nrow(holders)+(1:nrow(holders_t)))
		holders <- do.call(rbind,list(holders,holders_t))
		if(nrow(holders_t) < 1000) break()
		skip <- skip + 900
	}
	return(unique(holders))
}
holders_data <- list()
for(idx in 1:nrow(pools_data))
{
	holders_data[[idx]] <- get_holders(pools_data$Ticket_Address[idx],pools_data$Subgraph_Version[idx])
	message(idx)
}
saveRDS(holders_data,"data/holders_data.RDS")
```
</details>

Next, with the data successfully scraped, we read in the R data objects into our local environment and take a look at the raw data. To the rewards data, we add an indicator of the pool name for each.

```{r}
# Read the objects into the environment
pools_data <- readRDS("data/pools_data.RDS")
holders_data <- readRDS("data/holders_data.RDS")
rewards_data <- readRDS("data/rewards_data.RDS")

# Store the pool name int he rewards data
for (i in 1:length(rewards_data)) {
    if (nrow(rewards_data[[i]]) > 0) {
        rewards_data[[i]]$Pool <- pools_data$Name[i]
    }
}
```

We can take a look at the first few rows of the pool data:

```{r}
pools_data %>% head %>% kable
```

The holders of the first pool:

```{r}
holders_data[[1]] %>% head %>% kable
```

The past rewards for the first pool:

```{r}
rewards_data[[1]] %>% head %>% kable
```

With the data successfully scraped and parsed, we can begin our analysis.

## Analysis

To begin, we pre-process the data like so:

```{r}
## Flatten holders data
have_holders <- which(sapply(holders_data,nrow)>0)

h_names <- pools_data$Name[have_holders]
h_val <- pools_data$Ticket_Val_Base_Tok[have_holders]
h_valusd <- pools_data$Ticket_Val_USD[have_holders]
h_dec <- pools_data$Base_Tok_Decimals[have_holders]
h_holders <- holders_data[have_holders]

holders_flat <- do.call(rbind,mapply(function(x,y,z,w,xx)
    
cbind(x,Pool_Name=y,Value=z,USD=w,Dec=xx),h_holders,h_names,h_val,h_valusd,h_dec,SIMPLIFY=FALSE)) %>%
    as_tibble() %>%
    mutate(across(c(balance, Value, USD, Dec), as.numeric)) %>%
    mutate(Price = USD / Value) %>%
    mutate(Bal = balance / (10^Dec) * Price)
```

```{r}
ggplot(holders_flat, aes(x = Bal)) +
    geom_histogram(aes(fill = Pool_Name), bins=100) +
    scale_x_log10(labels = scales::dollar) +
    facet_wrap(~Pool_Name, scales = "free") +
    theme(
        axis.text.x = element_text(angle = 20, hjust = 1),
        legend.position = "off"
    )
```

```{r}
my_dollars <- function(x) {
    ifelse(x < .01, paste0("$", x), scales::dollar(x))
}

ggplot(holders_flat, aes(x = Bal)) +
    geom_histogram(aes(fill = Pool_Name), bins=100) +
    scale_x_log10(labels = my_dollars, breaks = c(1e-16, 1e-14, 1e-12, 1e-10, 1e-8, 1e-6, 1e-4, 1e-2, 1, 100, 10000, 1000000, 100000000)) +
    facet_wrap(~Pool_Name, scales = "free") +
    theme(
        axis.text.x = element_text(angle = 20, hjust = 1),
        legend.position = "off"
    )
```

```{r, fig.height=12}
ggplot(holders_flat, aes(x = Bal)) +
    geom_histogram(aes(fill = Pool_Name), bins=100) +
    scale_x_log10(labels = my_dollars, breaks = c(1e-16, 1e-14, 1e-12, 1e-10, 1e-8, 1e-6, 1e-4, 1e-2, 1, 100, 10000, 1000000, 100000000)) +
    facet_wrap(~Pool_Name, scales = "free_y", ncol = 1) +
    theme(
        axis.text.x = element_text(angle = 20, hjust = 1),
        legend.position = "off"
    )
```

### Average time between rewards

```{r}
rewards_data1 <- rewards_data %>%
    bind_rows() %>%
    as_tibble() %>%
    arrange(Awarded_Time) %>%
    mutate(TimeBetween = c(NA, diff(Awarded_Time)))

mean(rewards_data1$TimeBetween, na.rm = TRUE)
median(rewards_data1$TimeBetween, na.rm = TRUE)

ggplot(rewards_data1, aes(x = TimeBetween)) +
    geom_histogram()

ggplot(rewards_data1 %>% mutate(ID = 1:nrow(.)), aes(x = ID, y = TimeBetween)) +
    geom_point() +
    geom_line()
```

### How many in more than one pool?

```{r}
have_holders <- which(sapply(holders_data,nrow)>0)
h_names <- pools_data$Name[have_holders]
n_tokens <- pools_data$Base_Tok_Name[have_holders]
h_holders <- holders_data[have_holders]


flat <- do.call(rbind,mapply(function(x,y,z) cbind(x,Pool_Name=y, Token_Name=z),h_holders,h_names,n_tokens,SIMPLIFY=FALSE)) %>%
    as_tibble()

more_than_one <- flat %>%
    group_by(id) %>%
    summarise(Pools = length(unique(Pool_Name))) %>%
    filter(Pools > 1)

more_than_one %>% count()
```

### How many in exactly one pool?

```{r}
flat %>%
    group_by(id) %>%
    summarise(Pools = length(unique(Pool_Name))) %>%
    filter(Pools == 1) %>%
    count()
```

### Most pools

```{r}
flat %>%
    group_by(id) %>%
    summarise(Pools = length(unique(Pool_Name))) %>%
    arrange(desc(Pools))
```

### Computing Reward Diversity

```{r}
incidence <- flat %>% select(Address = id, Pool_Name) %>%
    filter(Address %in% more_than_one$id) %>%
    mutate(Count = 1) %>%
    spread(key = Pool_Name, value = Count)
incidence[is.na(incidence)] <- 0

test <- 1 - dist(incidence %>% select(-Address) %>% t, method = "binary", upper = TRUE, diag = TRUE) %>% as.matrix

test[1:5, 1:5]
``` 

```{r}
res <- sapply(1:nrow(rewards_data1), function(i) {
    x <- rewards_data1$Prize_External_ERC20[i]
    do.call(rbind, strsplit(gsub("\\[|\\]", "", strsplit(x, "][", fixed = TRUE)[[1]]), ";")) %>%
        as_tibble() %>%
        mutate(Pool = rewards_data1$Pool[i])
}) %>% bind_rows() %>%
    rename(Address = V1, Amount = V2, Name = V3, Ticker = V4) %>%
    mutate(Amount = readr::parse_number(Amount))

final <- res %>% group_by(Pool, Ticker) %>% summarise(Count = n()) %>% spread(key = Ticker, value = Count)
final[is.na(final)] <- 0

final %>% head %>% kable
```

