---
title: "Pool Together Analysis"
author: "Omni Analytics Group"
date: "8/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)

library(tidyverse)
library(knitr)
```

## Introduction

PoolTogether is a fun, open sourced and decentralized protocol for “no-loss” lotteries where users transfer their crypto into a shared pool that acts as a “prize-linked savings account”. These deposits double as lottery tickets that give the depositor a chance to win proportional to the amount of the pool they claim. Just like traditional savings accounts, all money in the pool earns a yield and each user can withdraw at their discretion.

At the time of this writing, \$3,801,856 worth of prizes have been awarded. Also, based on current deposits, the platform has been generating \$93,558 in interest per week for its users. Interested in earning? You can participate in PoolTogether on the Ethereum, xDai, Matic, and Binance Smart Chain networks.

True to the spirit of decentralization, the PoolTogether protocol is managed by holders of the $POOL governance token who draft and vote on proposals that control the prize creation tools and prize pools. Proposals are ratified through a voting process and are then implemented by the community.

## Data

```{r}
pools_data <- readRDS("~/Downloads/pools_data.RDS")
holders_data <- readRDS("~/Downloads/holders_data.RDS")
rewards_data <- readRDS("~/Downloads/rewards_data.RDS")
```

```{r}
pools_data %>% head %>% kable
```

```{r}
holders_data[[1]] %>% head %>% kable
```

```{r}
rewards_data[[1]] %>% head %>% kable
```

## Analysis

To begin, we pre-process the data like so:

```{r}
## Flatten holders data
have_holders <- which(sapply(holders_data,nrow)>0)

h_names <- pools_data$Name[have_holders]
h_val <- pools_data$Ticket_Val_Base_Tok[have_holders]
h_valusd <- pools_data$Ticket_Val_USD[have_holders]
h_dec <- pools_data$Base_Tok_Decimals[have_holders]
h_holders <- holders_data[have_holders]

holders_flat <- do.call(rbind,mapply(function(x,y,z,w,xx)
    
cbind(x,Pool_Name=y,Value=z,USD=w,Dec=xx),h_holders,h_names,h_val,h_valusd,h_dec,SIMPLIFY=FALSE)) %>%
    as_tibble() %>%
    mutate(across(c(balance, Value, USD, Dec), as.numeric)) %>%
    mutate(Price = USD / Value) %>%
    mutate(Bal = balance / (10^Dec) * Price)
```

```{r}
ggplot(holders_flat, aes(x = Bal)) +
    geom_histogram(aes(fill = Pool_Name), bins=100) +
    scale_x_log10(labels = scales::dollar) +
    facet_wrap(~Pool_Name, scales = "free") +
    theme(
        axis.text.x = element_text(angle = 20, hjust = 1),
        legend.position = "off"
    )
```

```{r}
my_dollars <- function(x) {
    ifelse(x < .01, paste0("$", x), scales::dollar(x))
}

ggplot(holders_flat, aes(x = Bal)) +
    geom_histogram(aes(fill = Pool_Name), bins=100) +
    scale_x_log10(labels = my_dollars, breaks = c(1e-16, 1e-14, 1e-12, 1e-10, 1e-8, 1e-6, 1e-4, 1e-2, 1, 100, 10000, 1000000, 100000000)) +
    facet_wrap(~Pool_Name, scales = "free") +
    theme(
        axis.text.x = element_text(angle = 20, hjust = 1),
        legend.position = "off"
    )
```

```{r, fig.height=12}
ggplot(holders_flat, aes(x = Bal)) +
    geom_histogram(aes(fill = Pool_Name), bins=100) +
    scale_x_log10(labels = my_dollars, breaks = c(1e-16, 1e-14, 1e-12, 1e-10, 1e-8, 1e-6, 1e-4, 1e-2, 1, 100, 10000, 1000000, 100000000)) +
    facet_wrap(~Pool_Name, scales = "free_y", ncol = 1) +
    theme(
        axis.text.x = element_text(angle = 20, hjust = 1),
        legend.position = "off"
    )
```

- Unique wallet deposits
- Pool Shared Holder Network Visualization
- Total rewards given out by token
- Token reward diversity by pool
- Time between rewards
- How participants only deposit into 1 pool?
- How many participants play in multiple pools?
- Who is the most “diversified” wallet addresses?
